SAM simétrico y asimétrico

### 2020-06-29 analysis/paper/paper.Rmd
draft figs

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

@@

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM."}
indexes %>% 
  .[lev == 700 & term != "full"] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  # geom_smooth(method = "lm", aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL)
```

```{r}
indexes_wide[, cor(asym, sym), by = lev] %>% 
  ggplot(aes(lev, V1)) +
  geom_line() +
  scale_x_level() + 
  scale_y_continuous(limits = c(0, 1)) +
  coord_flip()
```

```{r}

indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev)] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)]

indexes_2d %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
                # size = 0.3,
                # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ lev, labeller = labeller(term = lab_sam, lev = lev.lab))
```

Here is an example of inline code `r x` in the middle of a sentence.

>>>>

<!-- With the following code you can access and display values from the yml header above. -->
<!-- 
Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`
-->

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

@@

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM."}
indexes %>% 
  .[lev %in% c(30, 700) & term != "full"] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  # geom_smooth(method = "lm", aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL) +
  facet_wrap(lev~., labeller = labeller(lev = lev.lab), ncol = 1)
```

```{r cor-lev, fig.cap = "Correlation between the Symmetric and Asymmetric SAM at each level."}
indexes_wide[, cor(asym, sym), by = lev] %>% 
  ggplot(aes(lev, V1)) +
  geom_line() +
  scale_x_level() + 
  scale_y_continuous("Correlation", limits = c(0, 1)) + 
  coord_flip()
```

```{r cross-correlation, fig.cap = "Cross correlation between levels of the Full, Asymmetric and Symmetric SAM."}
indexes %>% 
  copy() %>% 
  .[, widyr::pairwise_cor(.SD, lev, time, estimate), by = .(term)] %>% 
  ggplot(aes(item1, item2)) +
  geom_contour_fill(aes(z = correlation), na.fill = 1) +
  scale_fill_divergent("Cross-correlation") +
  scale_x_level() +
  scale_y_level() +
  coord_equal() +
  facet_grid(. ~ term, labeller = labeller(term = lab_sam))
```

```{r compute-regr-pattern}
indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev)] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)] 
```

```{r 2d-regr, fig.cap = "Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef."}
indexes_2d %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ lev, labeller = labeller(term = lab_sam, lev = lev.lab))
```

```{r compute-regr-pattern-season}
indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev, season(time))] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)]

```

```{r 2d-regr-700, fig.cap = "Seasonal regression patterns of geopotential height at 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef."}
indexes_2d %>% 
  .[lat <= -20] %>% 
  .[lev == 700] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab))
```

```{r 2d-regr-30, fig.cap = "Seasonal regression patterns of geopotential height at 30 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef."}
indexes_2d %>% 
  .[lat <= -20] %>% 
  .[lev == 30] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab))
```

```{r wave-amplitude-700, fig.cap = "Planteray wave squared amplitudes for the regression patterns at 700 hPa."}
indexes_2d %>% 
  .[ lev == 700 & lat < -20] %>%
  .[, FitWave(estimate, 1:3), by = .(lat, lev, season, term)] %>% 
  ggplot(aes(lat, amplitude^2)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Squared amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~season, labeller = labeller(term = lab_sam)) 
```

```{r wave-amplitude-30, fig.cap = "Planteray wave squared amplitudes for the regression patterns at 30 hPa."}
indexes_2d %>% 
  .[ lev == 30 & lat < -20] %>%
  .[, FitWave(estimate, 1:3), by = .(lat, lev, season, term)] %>% 
  ggplot(aes(lat, amplitude^2)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Squared amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~season, labeller = labeller(term = lab_sam)) 
```

```{r compute-pp-regr} 
pp_regr <- CMAP() %>% 
  ReadNetCDF() %>% 
  normalise_coords() %>% 
  .[, precip := Anomaly(precip), by = .(lon, lat)] %>% 
  .[indexes_wide[lev == 700], on = .NATURAL] %>% 
  .[, FitLm(precip, asym, sym, se = TRUE, r2 = TRUE), by = .(lon, lat)] %>% 
  rm_intercept() %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(term)]
```

```{r regr-pp, fig.cap = "Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas."}
pp_regr %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), na.fill = TRUE, breaks = AnchorBreaks(0, 0.1, 0)) +
  geom_contour2(aes(z = p.value), breaks = 0.01) +
  geom_contour_fill(aes(z = as.numeric(is.na(estimate))), breaks = 1) +
  stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)), size = 0.4, alpha = 0.3) +
  geom_qmap() +
  scale_x_longitude() +
  scale_y_latitude() +
  scale_fill_divergent("Precipitation", guide = guide_colorstrip_bottom(), breaks = AnchorBreaks(0, 0.1, 0)) +
  coord_quickmap() +
  facet_wrap(~term, labeller = labeller(term = lab_sam))
```

```{r}
air_regr <- GISTEMP() %>% 
  ReadNetCDF(subset = list(time = range(indexes$time))) %>% 
  normalise_coords() %>% 
  # .[, air := Anomaly(air), by = .(lon, lat)] %>%
  .[indexes_wide[lev == 700], on = .NATURAL] %>% 
  .[, FitLm(air, asym, sym), by = .(lon, lat)] %>% 
  rm_intercept()
```

```{r,  fig.cap = "Regression pattern of surface temperature with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas."}
air_regr %>% 
  ggplot(aes(lon, lat)) +
  # geom_raster(aes(fill = estimate)) +
  geom_contour_fill(aes(z = estimate), na.fill = 0, breaks = AnchorBreaks(0, 0.1, 0)) +
  # geom_contour2(aes(z = p.value), breaks = 0.01) +
  stat_subset(aes(subset = is.na(estimate)), geom = "raster", fill = "gray50") +
  # stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)), size = 0.4, alpha = 0.3) +
  geom_qmap() +
  scale_x_longitude() +
  scale_y_latitude() +
  scale_fill_divergent("Surface temperature", limits = c(-0.5, 0.5), 
                       oob = scales::squish,
                       guide = guide_colorstrip_bottom(), breaks = AnchorBreaks(0, 0.1, 0)) +
  coord_quickmap() +
  facet_wrap(~term, labeller = labeller(term = lab_sam))
```
